Timeliness of security updates for Android

### 2015-03-25 securityupdates.tex
Rework Threats to Validity and destroy Discussion by blending into nearby sections

\section{Threats to validity}
\label{sec:validity}
\subsection{The \da\ data gives a conservative estimate of the Android version distribution}
\label{sec:representative}
\begin{figure}[!h]
 \centering
 \begin{subfigure}[b]{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/googleplayapi}
 \caption{Google Play data on proportion of devices running different API versions}
 \label{fig:play_api}
\end{subfigure}
\begin{subfigure}[b]{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
 \caption{\da\ data on proportion of devices running different API versions}
 \label{fig:da_api}
\end{subfigure}
\caption{Monthly Android API version data}
\end{figure}

The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
Unfortunately there is no ground truth of OS version information.
However Google has published API version information almost every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
We analysed the difference between the API version data from \da\ and Gooble Play, normalising for days since the API version was released.
This that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions (except for API version 17 which was particularly popular in a focused study discussed in the next section and so was temporarily overestimated when an old version).
This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
This allows us to have confidence in the OS version information.
\todolater{we want a statistical metric to claim this strongly with.}

\subsection{Changes in the \da\ sample}
\label{sec:da_changes}
The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app as there has been no advertising and few other attempts to increase usage.
However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular network operators.
One study was conducted in Norway with 654 installs and one in Bangladesh where 2463 users installed \da.
This is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
%We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 87.7\% rather than \daMeanInsecurityPercNominal, and the percentage of devices running the latest version to be 5.43\% rather than \daUpdatednessPercNominal.
Hence removing this anomaly would have no effect on our main conclusions.

Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.

\section{Discussion} \label{sec:discussion}
There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.

We have not included all the vulnerabilities from \avo. 
In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches which fix them. \url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
There also tends to be less public information available about manufacturer-specific vulnerabilities which makes them harder to tie down.

@@

We still do not know where new versions of Android come from.
Over time newer versions of Android come to dominate, but quantifying the proportion due to new phone purchases verses the proportion due to updates of existing devices is unknown.
Device Analyzer is the only available large data set we know of with longitudinal traces of Android OS version strings.
Unfortunately this data set only records \daNumUpdatesUpgrades\ upgrades which, broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month. This is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.

@@

>>>>

\section{Threats to validity}
\label{sec:validity}
\subsubsection{Comparing \da\ data with the ground truth}
\label{sec:representative}
\begin{figure}[!h]
 \centering
 \begin{subfigure}[b]{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/googleplayapi}
 \caption{Google Play data on proportion of devices running different API versions}
 \label{fig:play_api}
\end{subfigure}
\begin{subfigure}[b]{\columnwidth}
 \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
 \caption{\da\ data on proportion of devices running different API versions}
 \label{fig:da_api}
\end{subfigure}
\caption{Monthly Android API version data}
\end{figure}

The data from \da\ we used to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
Unfortunately there is no ground truth of OS version information.
However Google has published API version information almost every month since December 2009 and we have collated this information.\footnote{\url{http://androidvulnerabilities.org/play/historicplaydashboard}}
While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the \da\ data on API versions are similar to the Google data on API versions then the \da\ data on OS versions should be representative.
Figure~\ref{fig:play_api} shows the data from Google and Figure~\ref{fig:da_api} shows the data from \da\ and they appear similar.
We analysed the difference between the API version data from \da\ and Google Play, normalising for days since the API version was released.
This shows that the \da\ data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions (except for API version 17 which was particularly popular in a focused study discussed in the next section and so was temporarily overestimated when an old version).
This means that the OS version information from \da\ is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
Since most \da\ users are self selecting and install \da\ because they want to find out more about what their phone is doing or to aid research they may be biased and perhaps more likely to install updates.
This allows us to have confidence in the OS version information.
\todolater{we want a statistical metric to claim this strongly with.}

\subsubsection{The effect of focussed studies on \da}
\label{sec:da_changes}
The \da\ data is mostly generated by devices which have the \da\ app installed because their owner happened to come across the app as there has been no advertising and few other attempts to increase usage.
However two collaborations with network operators to install \da\ on customer devices resulted in large numbers of new \da\ users coming from particular network operators.
One study was conducted in Norway with 654 installs and one in Bangladesh where 2463 users installed \da.
This latter study is responsible for the sudden change in distribution of OS versions in \da\ beginning in August 2014 in Figures \ref{fig:norm_os}, \ref{fig:nvulnerabilities_heat} and \ref{fig:proportioninsecure} as for that period Bangladeshi users contributed over half of the \da\ data.
%We could normalise out the influence of the Bangladeshi users, however they are Android users too and they also need security updates.
To investigate the sensitivity of our results to the Bangladesh study we truncated our data before this study started and found the percentage of vulnerable devices to be 85.6\% rather than \daMeanInsecurityPercNominal, the percentage of devices running the latest version to be 5.71\% rather than \daUpdatednessPercNominal\ and the number of updates per year to be 1.48 rather than \daUpdatesPerYearNominal.
Hence removing this anomaly would have no effect on our main conclusions.

\subsubsection{The selection of vulnerabilities in \avo}
We have not included all the vulnerabilities from \avo. 
In particular, we have not included any device manufacturer specific vulnerabilities, even when these are widespread (such as vulnerabilities affecting Qualcomm\footnote{Qualcomm is particularly good at publicly disclosing vulnerabilities and the patches which fix them. \url{https://www.codeaurora.org/projects/security-advisories/}} chipsets) because it is difficult to work out which devices are affected.
There also tends to be less public information available about manufacturer-specific vulnerabilities which makes them harder to tie down.

@@

\subsubsection{Limitations}
We still do not know where new versions of Android come from.
Over time newer versions of Android come to dominate, but quantifying the proportion due to new phone purchases verses the proportion due to updates of existing devices is unknown.
Device Analyzer is the only available large data set we know of with longitudinal traces of Android OS version strings.
Unfortunately this data set only records \daNumUpdatesUpgrades\ upgrades which, broken down over \daOSMonthsOfData\ months and \daNumSigOSVersions\ OS versions is only \daUpdatesPerMonthPerVersion\ updates per version per month. This is not enough to build a statistically significant prediction of the expected transition to newer versions of Android.

@@

There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with and could be included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence Android.
